|

### 2021-11-23 07-Chp-Triangulation.Rmd
Working version after Chapter 7

This approach represents a significant step forward from the current evidence synthesis practice of synthesising evidence from only one approach (e.g. randomised controlled trials (RCTs), non-randomised studies of exposures (NRSE) or interventions (NRSI), etc.). The most common implementation of this method to date has been in the form of _qualitative triangulation_ - that is the identification and a narrative comparison, where the internal biases and differences between the study question and the causal question of interest (indirectness) of each result are discussed.

However, qualitative triangulation faces issues at scale. There is a need to include all evidence, as otherwise the potential for cherry-picking, yet as the number  as it is not possible to compare and contrast across multiple results in any meaningful way. This fact is also illustrated by previous exemplars of triangulation only considering three individual results. <!-- TODO CITATION NEEDED - original triangulation paper and Ference --> An alternative option previously employed is to assign bias to the output of a meta-analysis of similarly designed studies, but this loses information on the specific biases inherent to each results contributing to the summary estimate. For example, while it may be true that all non-randomised studies share some minimal level of bias, investigation of the biases applicable to each individual result gives useful additional information. Recent development in bias assessment have allowed for a much more granular consideration of the threats to the internal validity of a result. In combination, these limitations to qualitative triangulation detail a need for a more systematic way to integrate across multiple evidence sources as the number of individual results contributing to the exercise increases.

This chapter, in addition to providing a narrative comparison of the existing and new evidence identified in this thesis, builds on recent developments in risk-of-bias assessment and existing methods for bias-/indirectness-adjusted meta-analysis to propose a generalised framework for quantitative triangulation. The method, along with the methodological challenges to its implementation, is illustrated via a proof-of-concept case study examining the causal effect of average exposure to LDL-c at mid-life on Alzheimer's disease risk.  In the following section, I define important terms used throughout this chapter.

### Terminology {#tri-terminology}

<!-- Discuss and defined three things: -->

<!-- * Qualitative vs quantitative triangulation -->
<!-- * Bias vs. indirectness -->
<!-- * Additive vs proportion effect of bias/indirectness -->

For the sake of clarity, and given the multiple competiting defitions used to define key concepts on this topic, I begin by defining the terms as used in this chapter.

<!-- TODO Does Julian have a good reference for the conflict between different terms used for biases? -->

__Bias vs indirectness__

Both

Bias (or internal bias) in the result is defined as systematic deviation from result the "idealised" version of the study (that is, )

In contrast, the indirectness of the result (also called generalisability or applicability) is defined here as the discrepancy between the research question addressed by the "idealised" study and the causal question of interest. Problems with indirectness can be arise from differences in population, intervention/exposure, outcomes and

Of note, the original method employed in this chapter refers indirectness as "external bias" - however. This distinction is important, as bias and indirectness describe two different issues, and referring to both as bias

In addition, indirectness is the widely used term, including by widely-adopted methods such as the GRADE framework.[@guyatt2011]

__Additive vs proportional effects__

Additive biases (identified by a "Favours experimental"/"Favours comparator" response) are XXXX. A key example of additive bias is immortal time bias, which will always favour the intervention.

Proportional biases (identified by a "Towards null"/"Away from null" response) are those related to the magnitude of the effect. A key example of this type of bias is non-differential misclassification bias, which will bias the effect estimate towards the null.

The effects of both bias and indirectness can be either additive or propotional.

Additive bias/indirectness is

Both bias and indirectness can be defined as additive or propotional. Additive bias/indirectness is that which can shift the p

@@

This triangulation exercise draws on the research produced in each of the preceding chapters. More specifically, this chapter builds on the comprehensive systematic review presented in chapter \@ref(sys-rev-heading). It also incorporates the new evidence produced by the analysis of the association of statin use with dementia outcomes in the CPRD (Chapter \@ref(cprd-heading)) and of the association of lipids with dementia outcomes in previously unanalysed datasets accesssed via the DPUK (Chapter \@ref(ipd-heading)).

Table \@ref(tab:thesisOverview-table) summarises the research each approach has attempted to address, the exposures and outcomes considered in each study, and the contribution of each chapter to the triangulation exercise presented here.

\blandscape
<!----------------------------------------------------------------------------->
(ref:thesisOverview-caption) Summary of studies included in this thesis, and used as evidence sources in this triangulation exercise. Note, Chapter \@ref(sys-rev-tools-heading) is intentionally not included in this table, as it describes a research tool rather than a research study.

@@

The risk of bias tools used to assess each result are described in detail in Section \@ref(risk-of-bias). In summary, RTCS were assessed using the RoB 2 tool;[@sterne2019] NRSI  using the ROBINS-I tool;[@sterne2016] NRSE using the ROBINS-E tool (see Section \@ref(#rob-tools-nrse));[@morganr2020; @french2019] and MR studies using the Mamluk et al. tool.[@mamluk2020]

<!-- TODO Talk about how the CPRD and IPD analysis results were assessed here! -->

<!-- Risk of bias assessment was performed for each of the included cohorts using the risk-of-bias assessment tool for non-randomised studies of exposures introduced previously (for a more detailed discussion of this tool, see Section \@ref(rob-tools-nrse)). -->

<!-- Due to the inherent conflict of interest in assessing my own analysis, the result were also assessed by two independent reviewers. -->

### Qualitative triangulation

As discussed in the introduction to this Chapter (Section \@ref(triangulation-overview)), as part of a narrative synthesis of the evidence all information sources were grouped by outcome, and systematically compared and contrasted. Potential reasons for heterogeneity between study designs were examined with specific reference to the risk-of-bias assessments performed.

<!-- TODO Use the fact that it doesn't make sense to qualitatively compare at the single result level to highlight the need for a more quantitative approach - usually default to summary at the -->

<!-- IDEA Note the text by Turner on the fact that you can just adjust for a single type of bias - a good example of this would be to try and adjust for internal biases so you get the best possible estimate and then apply it to something like a assumed life-course model - steal from NIHR text, and refer to the appendix  -->

This analysis built on the detailed risk of bias assessments performed and presented as part of the systematic review (Chapter \@ref()/\@ref()). Similarly, risk-of-bias assessments were performed for each new source of evidence presented in this thesis, details of which are available in Appendix \@ref(?).

<!-- TODO Need to actually do these, and include results in the analysis. -->

@@

In addition to the qualitative discussion of evidence, I attempted to integrate the numerical results of the multiple approaches. This proof-of-concept approach incorporates recent advancements in the way that bias in results is assessed (move from checklists to domain-based assessment tools)[@sterne2019] and existing methods for bias-/indirectness-adjusted meta-analysis[@turner2009] to to illustrate how causal questions could be addressed under this quantitative triangulation framework.

This proposed framework involves several steps, as described in detail in the following sections. In summary these steps are:

1. Define the causal question of interest
1. Identify of relevant evidence sources and standardisation of effect directions
1. Assess and visualise the extent and direction of bias in each result
1. Define modifying terms for bias and indirectness in each result
1. Caclulate bias-/indirectness-adjusted results and perform meta-analysis

In order to illustrate the bias-/indirectness adjustment process detailed in the subsequent sections, the process for calculating the adjusted estimate for a single study is described in detail for a single study. <!-- TODO CITATION NEEDED need to cite single paper here --> Following this, the process for an example causal question as a case study, as defined below.

@@

Given that this quantitative analysis is intended as a proof-of-concept for the framework, rather than aiming to produce findings that should guide clinical practice, I consider only one of the multiple possible causal questions related to the broad topic of the effect of lipids on dementia outcomes.

Using the ROBINS-E framework I defined the parameters of my causal question of interest:

*	_Population of interest_: General population
*	_Exposure of interest_: Low density lipoprotein cholesterol
*	_Exposure window of interest_: Mid-life (45-60)
* _Summary measure of exposure over time_: Average exposure

This question was chosen for a number of reason. Firstly, given the long prodomal period between the onset of physiological changes in the brain and presentation of dementia symptoms, it seems likely that conditions during the mid-life period is particularly important. Secondly, examining the causal impact of lipids on a specific outcome (in this case, Alzheimer's disease), rather than a composite outcome (all-cause dementia), focuses on an single causal pathway rather than including conditions that likely have very different mechanisms of disease. Finally, and relevant to its use as a case study for the quantitative triangulation approach, this question had several relevant sources of evidence available (versus, for example, any question looking at vascular dementia).

<!-- TODO This table will describe the causal questions of interest, following the ROBINS-E framework as columns, and list the included studies adn the relative weights of each. -->

@@

####	Identify relevant studies

Once the causal question of interest had been defined, relevant individual results were obtained from the data sources described in Section \@ref(tri-data-sources). In a broad sense, this meant pulling out studies that examined the relationship between LDL-c and Alzheimer's disease risk, either directly (non-randomised studies of LDL-c levels) or indirectly (non-randomised studies of statins, Mendelian randomisation studies using genetic instruments for LDL-c levels).

Once the set of relevant results were identified, effects were standardised to refer to the risk of Alzheimer's disease resulting from a "reduction" in LDL-c. For studies of statins, th. However, for both non-randomised studies and Mendelian randomisation analysis of lipid levels, the effect is usually reported per 1-SD increase in LDL-c. In this case, the effect estimates were inverted to ensure consistency across study designs.

@@

For the sake of consistency, I mapped the different acceptable judgements in each risk-of-bias tool to a harmonised set of judgements: "Low", "Moderate","High". The exact mapping can be seen in Table \@ref(tab:robLevelsMapping-table). Of note, no mapping was performed for the critical risk-of-bias levels present in the tools for non-randomised studies. This is because current best practice in evidence synthesis is to exclude all studies at critical risk of bias for further quantitative synthesis.[@sterne2016]

&nbsp;<!----------------------------------------------------------------------->  
(ref:robLevelsMapping-caption) robLevelsMapping

(ref:robLevelsMapping-scaption) robLevelsMapping

@@

In addition to assessing the extent of bias in each domain, as defined by the three levels detailed above, I attempted to record the predicted direction and type of bias. This was achieved via an additional question at the end of each bias domain, which asked assessors to predict the expected direction of bias in that domain. Note that this question is currently only formally included in two existing tools (ROB2/ROBINS-I), but I employed an identical approach during assessment of non-randomised studies of exposure and Mendelian randomisation studies. Acceptable responses to this question included:

@@

As indicated in the options above, the response to this question is also used to determine the predicted type of bias (additive or proportional - see Section \@ref(tri-terminology)) in addition to the direction. For additive biases, determining the absolute direction of bias was simply a case of whether the bias shift the effect estimate to the left or right on an imaginary forest plot. For example, if a bias was predicted to favour the intervention, then the estimate would be shifted to the right.

In contrast, for proportional biases, whether the estimate should be increased or decreased proportionally depends on the current position of the point estimate relative to the null. For example, if the effect estimate represents a protective effect (below the null), then bias towards the null would be adjusted for by moving the effect estimate propotionally to the right. In contrast, if the effect of the intervention is harmful (effect estimate above the null), bias towards the null would be adjusted for by moving the effect estimate propotionally to the left.

@@

(ref:exampleDirection-cap) Illustration of calculating the absolute direction of biases (indicated by arrows), based on bias type. For additive biases, the direction is consistent regardless of the position of the effect estimate relative to the null. In contrast, the position of the effect estimate is accounted for when determining the absolute direction of proportional biases.

@@

Note that for some domains, only one type of bias is allowed. For example, in the confounding domains in the ROBINS-I/ROBINS-E tools, only the "Favours experimental"/"Favours comparator", and "Unpredictable" options are available, meaning that bias in this domain will always be considered additive. Additionally, the predicted direction and type of bias were not recorded for domains at "Low" risk of bias.

@@

To aid with the triangulation exercise and to act as a sense-check, I created a new method to visualise the extent, direction and type of bias in each result to enable detailed comparison across different studies contributing to a synthesis. Similar to the standard paired forest plots introduced in Chapter \@ref(sys-rev-), these "bias direction" plots show the  level of bias across the domains in each study is reported using coloured blocks. However, in contrast to the previous plots, symbols are used to illustrate the predicted absolute direction of bias in that domain, as discussed in the previous section. Distinct symbol pairs are used to denote additive and proportional biases.

Convience functions in the `triangulate` package (see Section \@ref(tri-software)) were used to preprocess risk of bias data for use with both bias-direction plots and adjusted meta-analyses, as discussed in subsequent sections. These bias direction plots themselves were created using the risk-of-bias visualisation tool described in Appendix \ref(?).

@@

@@

Reasonable values for the position and spread of the prior distributions were informed by a previous reported expert bias elicitation exercise.[@turner2009] Using open data from that study, I calculated the mean and maximum values for the mean and variance of the adjustment distributions across all biases considered by that exercise (see Appendix \@ref(?) for more details on how these were calculated). For the purposes of this illustrative analysis and to highlight the generalised nature of the code used to perform the analysis (see Section \@ref(tri-software)), I used the calculated values to inform two sets of modifying distributions (Table \@ref(tab:priorsAdd-table)) for additive biases, and a single set for proportional biases. In the first scenario, the relationship between the position assigned to each additive judgement is linear. In the second scenario, the step from "moderate" to high" is twice that from "low" to "moderate". Across both additive and proportional biases, the variance of the "high" risk--of-bias judgement distribution was defined to be greater than that for the "moderate" risk-of-bias judgement.

@@

(ref:priorsAdd-scaption) Additive bias distributions

@@

Where the direction of bias in a domain was unpredictable for a given result, the mean of the adjustment distribution for that domain was set to 0 but the appropriate variance for the recorded extent of bias was retained (e.g. for an unpredictable "high" assessment, the distribution would be $N(0,0.12)$). In other words, adjusting for bias in a domain with an unpredictable direction of bias has no impact on the position of the effect estimate, but does increase the uncertainty around it.

@@

####	Assessing and assigning modifying values to indirectness

An identical approach was employed to assess and assign prior distributions to the risk of indirectness in each result. The target question for each result was compared against the causal question of interest with respect to three domains (population, intervention/exposure, outcome), <!-- TODO CITATION NEEDED --> again using the scale of "Low/"Moderate"/"High" to quantify the extent of indirectness. All indirectness were defined _a priori_ as being proportional in nature, in line with previous work on this topic.[@turner2009]

As crude illustrative example, consider the causal question of the interest in this exercise, the effect of LDL-c in mid-life on Alzheimer's disease. Here, studies of lipids levels in this time period would require minimal adjustment, whereas studies examining statin use (indirect exposure/intervention) in late-life (indirect population) would be downweighted due to reduced relevance to the causal question.

@@

Using the the method reported in Turner _et al._,[@turner2009] total additive and proportional bias and indirectness were calculated for each result. Briefly,

<!-- TODO Talk about how total add/prop biases calculated -->

Once the total additive and proportional bias and indirectness for each result were computed, they were used to define an adjusted estimate for each result included in the synthesis.

Using the method defined in Turner _et al_,[@turner2009] the adjusted estimate for each result, $\hat\theta$, is defined as:

@@

@@

Once each individual result has been adjusted, I performed a random effects meta-analysis using the unadjusted and adjusted (under the two scenarios defined above) results. <!-- TODO CITATION NEEDED -->

<!-- TODO Deal with the fact that the Panel wasn't super useful -->

@@

### Qualitative triangulation

@@

@@

<!-- TODO Need to talk about fibrates here too -->

__Statins and all-cause dementia__

This conflicts with the findings of our analysis, where statin use was associated with an increased risk of all-cause dementia (`r anydem_text`). Some of the included studies in the meta-analysis specifically exclude vascular dementia from the definition of all-cause dementia,[@chao2015] which may limit the ability for comparison with our findings for the all-cause dementia outcome.

Additionally, a previous analysis of the THIN EHR database using a propensity-score matched analysis found a protective effect of statins on all-cause dementia (HR:0.81, 95%CI:0.69-0.96)[@smeeth2009].

@@

__Statins and Alzheimer's disease__

Our results are broadly in line with the findings of two distinct approaches examining the effect of statin treatment on subsequent Alzheimer's disease. No randomized trials of statins for the prevention of Alzheimer's disease have been reported, but a recent meta-analysis of 20 observational studies found statins were associated with a reduced risk of Alzheimer's disease (RR 0.69, 95% CI 0.60–0.80) with stronger evidence than observed in our analysis.[@poly2020] This review included case-control studies and analyses likely to be at risk of immortal time bias, which may account for the discrepancy with our findings. Additionally, a recent Mendelian randomization study examining the effect of genetic inhibition of HMGCR on Alzheimer’s disease (a genetic proxy for statin treatment) provided equivocal evidence (OR: 0.91, 95%CI: 0.63-1.31) but was consistent with our results.[@williams2020]

Our additional analyses stratified by statin properties found little evidence of differences in associations of lipophilic and hydrophilic statins and incidence of Alzheimer’s disease, consistent with a recent meta-analysis of observational studies.[@chu2018]

@@

__Statins and vascular/other dementia__

Far fewer studies have tested the association between lipid-regulating agents and vascular dementia or other dementias. A recent review found four observational studies examining the association of statins and vascular dementia found limited evidence for an effect (RR:0.93, 95% CI 0.74–1.16).[@poly2020] This contrasts with the harmful association found in our analysis (`r vasdem_text`). When stratifying by lipid properties, lipophilic statins were more harmful than hydrophilic statins in vascular dementia, potentially due to their ability to cross the blood brain barrier.

__Other drug classes__

Apart from statins, few studies examining a lipid-regulating agent have been reported (Chapter \@ref(sys-rev-methods-heading)/\@ref(sys-rev-results-heading)).

One of the few classes for which a evidence was available were fibrates, which found little evidence of an association with all-cause dementia was identified,[@ancelin2012] inconsistent with our finding that patients prescribed fibrates had higher all-cause dementia risk than those prescribed other lipid lowering agents.

A previous Mendelian randomization study found little evidence that genetic variants that proxy for ezetimibe affect risk of Alzheimer’s disease (OR: 1.17, 95%CI: 0.73-1.87),[@williams2020] consistent with our findings. Note that this study was published in 2020 and so was not included in the review.

@@

@@

#### Case study: effect of LDL-c at midlife on Alzheimer's disease

>>>>

This approach represents a significant step forward from the current practice of synthesising evidence from only one type of study (e.g. randomised controlled trials (RCTs), non-randomised studies of exposures (NRSE) or interventions (NRSI)). The most common implementation of this method to date has been in the form of _qualitative triangulation_ - that is the identification and narrative comparison of diverse evidence sources with respect to their underlying bias structures.[@lawlor2016]

However, qualitative triangulation faces issues at scale. There is a need to include all evidence to avoid potential confirmation bias,[@dubroff2018] yet as the number of evidence sources increases, it is not possible to narratively compare and contrast across multiple results in any meaningful way. This fact is illustrated by previous exemplars of triangulation only considering a small number of individual results.[@ference2014; @lawlor2016] To overcome this limitation, some attempts have considered the output of a meta-analysis of similarly-designed studies to be a single source of evidence, and have assessed bias in relation to this summary result. While this approach keeps the number of individual sources of evidence manageable, it loses useful information on the specific biases inherent to each result contributing to the summary estimate. For example, while it may be true that all non-randomised studies share some minimal level of bias, competing extents and directions of bias in individual results may be masked by the use of summary estimates. In addition, previous attempts at qualitative triangulation have not systematically assessed the indirectness of the results, that is the differences between the question of interest and the question addressed by a study.

Taken together, these limitations to qualitative triangulation detail a need for a more systematic way to integrate across multiple evidence sources as the number of individual results contributing to the exercise increases. One such approach is _quantitative triangulation_, where results from different evidence sources are combined numerically while accounting for the key sources of bias and indirectness in each. This chapter, in addition to providing a narrative comparison of the existing and new evidence identified by this thesis, builds on recent developments in risk-of-bias assessment and methods for bias-/indirectness-adjusted meta-analysis to propose a generalised framework for quantitative triangulation. The framework, along with the methodological challenges to its implementation, is illustrated via two case studies examining the causal effect of lipids on dementia outcomes.

@@

This triangulation exercise draws on the research produced in each of the preceding chapters. More specifically, this chapter builds on the comprehensive systematic review of existing evidence presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). This evidence base is then supplemented by new evidence on the association of statin use with dementia outcomes in the CPRD (Chapter \@ref(cprd-heading)) and the association of lipids with dementia outcomes in previously unanalysed datasets accessed via the DPUK (Chapter \@ref(ipd-heading)).

Table \@ref(tab:thesisOverview-table) summarises the research each data source has attempted to address, the exposures and outcomes considered in each case, and the contribution of each chapter to the triangulation exercise presented here.

\blandscape
<!----------------------------------------------------------------------------->
(ref:thesisOverview-caption) Summary of studies included in this thesis and used as evidence sources in this triangulation exercise. Note, Chapter \@ref(sys-rev-tools-heading) is intentionally not included in this table, as it describes a research tool rather than a research study.

@@

### Qualitative triangulation {#qual-tri}

As discussed in the introduction to this Chapter (Section \@ref(triangulation-overview)), as part of a narrative synthesis of the evidence all information sources were initally grouped by outcome, and the findings from each source were narratively compared and contrasted. Potential reasons for heterogeneity between study designs were examined with specific reference to the risk-of-bias assessments performed.

This analysis built on the detailed risk-of-bias assessments performed and presented as part of the systematic review (Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading)). Similarly, risk-of-bias assessments were performed for each new source of evidence presented in this thesis. The risk of bias tools used to assess each result are described in detail in Section \@ref(risk-of-bias), but in summary, RCTS were assessed using the RoB 2 tool;[@sterne2019] NRSI using the ROBINS-I tool;[@sterne2016] NRSE using the ROBINS-E tool (see Section \@ref(rob-tools-nrse));[@morganr2020; @french2019] and MR studies using the Mamluk et al. tool.[@mamluk2020]

@@

In addition to a qualitative discussion of the evidence, I attempted to integrate the numerical results of the multiple approaches under a proposed quantitative triangulation framework. This approach incorporates recent advancements in the way that bias in results is assessed (namely the move to domain-based assessment tools)[@sterne2019] and existing methods for bias-/indirectness-adjusted meta-analysis[@turner2009] to illustrate how causal questions could be addressed under this framework.

This proposed framework involves several steps, described in detail in the following sections. In summary, these steps are:

1. Define the causal question of interest
1. Identify of relevant evidence sources and standardise effect directions
1. Assess and visualise the extent and direction of bias in each result
1. Define modifying terms for bias and indirectness in each result
1. Calculate bias-/indirectness-adjusted results and perform meta-analysis

In order to illustrate the bias-/indirectness adjustment process detailed in the subsequent sections, the process for calculating the adjusted estimate is described in detail for a single result. Following this, the framework is applied to two case studies, as defined below.

@@

Using the ROBINS-E framework,[@morganr2020] I defined the parameters of my causal questions of interest:

__Case study #1: Effect of mid-life LDL-c on Alzheimer's disease__

*	_Population of interest_: General population
*	_Exposure of interest_: Low density lipoprotein cholesterol
*	_Exposure window of interest_: Mid-life (45-60)
* _Summary measure of exposure over time_: Average exposure 
* _Outcome of interest:_ Alzheimer's disease

__Case study #2: Effect of mid-life triglycerides on vascular dementia__

*	_Population of interest_: General population
*	_Exposure of interest_: Triglycerides
*	_Exposure window of interest_: Mid-life (45-60)
* _Summary measure of exposure over time_: Average exposure
* _Outcome of interest:_ Vascular dementia

&nbsp;

These questions were chosen for several reasons. Firstly, work presented in previous chapters had identified some evidence for an association in these exposure/outcome pairs. Secondly, given the long prodomal period between the onset of physiological changes in the brain and presentation of dementia symptoms, it seems likely that conditions during the mid-life period are particularly important. Finally, examining the causal impact of a lipid faction on a specific outcome (e.g. Alzheimer's disease), rather than a composite outcome (e.g. all-cause dementia), focuses on an single causal pathway rather than including several conditions in the outcome definition that likely have very different mechanisms of disease. In addition, previous work presented in this thesis illustrated that the component conditions (in this case Alzheimer's disease and vascular dementia) are likely to be subject to very different confounding structures (see Section \@ref(hold)), and use of an composite all-cause outcome in the triangulation framework may mask this.

@@

####	Identify relevant results

Once the causal question of interest had been defined, relevant individual results were obtained from the data sources described in Section \@ref(tri-data-sources). In a broad sense, this meant extracting results that examined the relationship between the exposure/outcome pair, either directly (e.g. non-randomised studies of LDL-c levels) or indirectly (e.g. non-randomised studies of statins, or Mendelian randomisation studies using genetic instruments for LDL-c levels).

Once the set of relevant results were identified, effects were standardised to refer to the risk resulting from a "reduction" in the lipid fraction. For both non-randomised studies and Mendelian randomisation analysis of lipid levels, the association is usually reported per 1-SD increase in the lipid fraction. In this case, the effect estimates were inverted to ensure consistency across study designs.

@@

Risk of bias was assessed for each result as discussed in Section \@ref(qual-tri). For the sake of consistency and ease of computing, I mapped the different acceptable risk-of-bias judgements in each tool to a harmonised set: "Low", "Moderate", and "High". The exact mapping can be seen in Table \@ref(tab:robLevelsMapping-table). Of note, no mapping was performed for the critical risk-of-bias levels present in the tools for non-randomised studies. This is because current best practice in evidence synthesis is to exclude all studies at critical risk of bias for further quantitative synthesis.[@sterne2016]

&nbsp;<!----------------------------------------------------------------------->  
(ref:robLevelsMapping-caption) Mapping of the different acceptable levels of bias across the assessment tools to a harmonised set. Note that no mapping for "Critical" risk of bias was performed, as best practice reccommends that studies at critical risk of bias are excluded from quantitative syntheses.

(ref:robLevelsMapping-scaption) Mapping of risk-of-bias judgements across assessment tools

@@

In addition to assessing the extent of bias, as defined by the three levels detailed above, I attempted to record the predicted direction and type of bias in each domain. This was achieved via an additional question at the end of each bias domain, which asked assessors to predict the expected direction of bias in that domain. Note that this question is currently only formally included in two existing tools (ROB2/ROBINS-I), but I employed an identical approach during assessment of non-randomised studies of exposure and Mendelian randomisation studies. Acceptable responses to this question included:

@@

As indicated in the options above, the response to this question is also used to determine the predicted type of bias as either additive or proportional. Proportional biases are dependent on the magnitude of the effect, while additive biases are not.[@turner2009]

For additive biases, determining the absolute direction of bias was simply a case of whether the bias is expected to shift the effect estimate to the left or right on an imaginary forest plot. For example, the estimate would be shifted to the right if a bias was predicted to favour the comparator (Figure \@ref(fig:exampleDirection)).

In contrast, for proportional biases, whether the estimate should be increased or decreased proportionally depends on the current position of the point estimate relative to the null. For example, if the effect estimate represents a protective effect (below the null), then bias towards the null would be adjusted for by moving the effect estimate proportionally to the right. In contrast, if the effect of the intervention is harmful (effect estimate above the null), bias towards the null would be adjusted for by moving the effect estimate proportionally to the left. Both scenarios are illustrated using example data in Figure \@ref(fig:exampleDirection).

@@

(ref:exampleDirection-cap) Illustration of calculating the absolute direction of biases (indicated by arrows), based on bias type. For additive biases, the direction is consistent regardless of the position of the effect estimate relative to the null - bias "favouring the comparator" will always pull the effect estimate to the right. In contrast, the position of the effect estimate is accounted for when determining the absolute direction of proportional biases, defined as towards or away from the null.

@@

The predicted direction and type of bias were not recorded for domains at "Low" risk of bias. Finally, for some domains, only one type of bias is allowed. For example, in the confounding domains in the ROBINS-I/ROBINS-E tools, only the "Favours experimental"/"Favours comparator" and "Unpredictable" options are available, meaning that bias in this domain will always be considered additive.

@@

To aid with the quantitative triangulation exercise, I created a new method to visualise the extent, direction and type of bias in each result, enabling detailed comparison across different studies contributing to a synthesis. Similar to the standard paired forest plots introduced in Chapter \@ref(sys-rev-), these "bias direction" plots show the level (or extent) of bias across the domains for each result reported using coloured blocks. However, in contrast to the previous forest plots, symbols are used to illustrate the predicted absolute direction of bias in that domain. Distinct symbol pairs are used to denote additive and proportional biases.

@@

<!-- QUESTION sign is negaitve if to the left, because you are taking it away in the final formula and it therefore needs a negative sign to be positive overall! -->

@@

Reasonable values for the position and spread of the prior distributions were informed by a previously-reported expert bias elicitation exercise.[@turner2009] Using open data from that study, I calculated the mean values for the mean and variance of the adjustment distributions across all biases considered by that exercise (see Appendix \@ref(?) for more details on how these were calculated). For the purposes of this illustrative analysis and to highlight the generalised nature of the code used to perform the analysis (see Section \@ref(tri-software)), I used the calculated values to inform two scenarios of modifying distributions (Table \@ref(tab:priorsAdd-table)) for additive biases, and a single set for proportional biases. In the first scenario, the relationship between the position assigned to each additive judgement is linear. In the second scenario, the step from "Moderate" to "High" is twice that from "Low" to "Moderate". Across both additive and proportional biases, the variance of the "High" judgement distribution was defined to be greater than that of the "Moderate" judgement.

@@

(ref:priorsAdd-scaption) Prior distributions mapped to different extents of bias.

@@

Where the direction of bias in a domain was unpredictable for a given result, the mean of the prior distribution for that domain was set to 0 but the appropriate variance for the recorded extent of bias was retained (e.g. for an unpredictable "High" additive bias, the distribution would be $N(0,0.1)$ under Scenario 1). In other words, adjusting for bias in a domain with an unpredictable direction of bias has no impact on the position of the effect estimate but does increase the uncertainty around it.

@@

####	Assessing and assigning prior values to indirectness

The indirectness of a result (also termed external bias, relevance, generalisability, or applicability) is defined here as the discrepancy between the research question addressed by the "idealised" study and the causal question of interest.

An identical approach was employed to assess, and assign prior distributions to, the indirectness in each result. The target question for each result was compared against the causal question of interest with respect to three domains, defined as important by the GRADE framework for assessing the certainty of evidence:[@guyatt2011] population, intervention/exposure, and outcome. I again used the scale of "Low/"Moderate"/"High" to quantify the extent of indirectness in each domain. All indirectness was defined _a priori_ as being proportional in nature (i.e. depending on the magnitude of the effect), in line with previous work on this topic.[@turner2009;@thompson2011] As above, data from the previous elicitation exercise were used to inform reasonable prior distributions for each level of indirectness (Table \@ref(tab:priorsIndirect-table)).

&nbsp;

<!----------------------------------------------------------------------------->
(ref:priorsIndirect-caption) Prior distributions for extent of indirectness, on the log scale.

(ref:priorsIndirect-scaption) Prior distributions mapped to different extents of indirectness.

```{r priorsIndirect-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

&nbsp;

As an illustrative example, consider the first causal question of the interest in this exercise, the effect of LDL-c in mid-life on Alzheimer's disease. Here, studies of lipids levels in this time period would require minimal adjustment, whereas studies examining statin use (indirect exposure/intervention) in late-life (indirect population) would be down-weighted due to reduced relevance to the causal question.

@@

Using the the method reported in Turner _et al._,[@turner2009] total additive and proportional bias and indirectness were calculated and used to define an adjusted estimate for each result included in the synthesis.

The adjusted estimate for each result, $\hat\theta$, is defined as:

@@

where $\mu_{i\delta}^{\mathrm{B}}$ and $\mu_{i\beta}^{\mathrm{B}}$ refer to the total additive and proportional bias, and  $\mu_{i\delta}^{\mathrm{In}}$ and $\mu_{i\beta}^{\mathrm{In}}$  refer to the total additive and proportional indirectness. Note that in this exercise, the total additive indirectness $\mu_{i\beta}^{\mathrm{In}} = 1$, as all indirectness was defined _a priori_ as being proportional.

@@

Once each individual result had been adjusted using this approach, I performed a random effects meta-analysis using the unadjusted and adjusted results. I also compared the adjusted results under the two scenarios of additive bias defined. results. I extracted the overall effect estimates, along with measures of heterogeneity in each case ($\tau^2$, $I^2$).[@higgins2003; @higgins2008b]

<!-- IDEA Sensitivity analysis - set means to zero and only adjust for variance. Essentially adding uncertainty based on the amount of bias but not moving the effect estimate -->

@@

### Qualitative triangulation / narrative comparison

__[Question: the "strength of evidence" language here makes for fairly ambiguous reading - e.g. "very weak evidence". Should I also include the effect estimates from each approach, which will help to clarify, if it is a narrative synthesis?]__

@@

I identified no consistent effect of any lipid fraction on any dementia outcome  across the evidence sources presented in this thesis, through there was some suggestion of a protective effect of LDL-c lowering (via statin use) on all-cause dementia and Alzheimer's disease, and a harmful effect of raised triglyceride levels on vascular dementia.

@@

Analysis of lipid levels across both the systematic review (Chapter \@ref(hold)) and the IPD analysis (Chapter \@ref(hold)) both found extremely weak evidence for an association of any lipid fraction with all-cause dementia. Similarly, both a meta-analysis of RCTs of statin use and of MR analyses of genetic instruments that mimic statin use found very weak evidence of an association, providing indirect evidence on the effect of LDL-c lowering on this outcome.

In contrast, the meta-analysis of NRSI of statin use found a slight protective effect, though many of the studies in this analysis were found to be at risk of immortal time bias, which would be expected to favour the intervention. Finally, analysis of the association of statin use with all-cause dementia in the CPRD (Chapter \@ref(hold)) found evidence of a harmful effect, though this is likely to be driven by the substantial confounding by indication identified in the vascular dementia subgroup.

For fibrates, an alternative lipid-regulating agent, there was some evidence of an increased risk of all-cause dementia in the CPRD, though this did not replicate in the meta-analysis of previous studies examining the association of fibrates and all-cause dementia.

@@

Meta-analysis of lipid levels in the systematic review (Chapter \@ref(hold)) found extremely weak evidence for an association of any lipid fraction with Alzheimer's disease. Similarly, a meta-analysis of MR analyses of genetic instruments that mimic statin use found very weak evidence of an association.

In contrast, the meta-analysis of NRSI of statin use found a slight protective effect, though similar to the all-cause dementia outcome, many of the studies in this analysis were found to be at risk of immortal time bias. However, the CPRD analysis (Chapter \@ref(hold)) attempted to account for this potential bias through the use of time-varying treatment indicator. It found very weak evidence for an effect, though the potential for differential misclassification on the basis of the exposure limit the validity of this finding.

The IPD analysis (Chapter\@ref(hold)) provided no information on this outcome, as the sole cohort containing data on Alzheimer's disease (Whitehall II) had previously been analysed in respect to this outcome and so was included in the meta-analysis of NRSE.

@@

There was a general absence of existing evidence on vascular dementia outcomes.  Meta-analysis of the small number of studies examing  lipid levels in the systematic review (Chapter \@ref(hold)) found extremely weak evidence for an association of any lipid fraction with vascular dementia. Similarly, a meta-analysis of statin use found very weak evidence for an effect on this outcome. The CPRD analysis (Chapter \@ref(hold)) found evidence for a harmful association of statin use with vascular dementia, though this finding is likely to be the result of severe confounding by indication, as identified through the use of control outcomes.

Finally, the IPD analysis (Chapter \@Ref(hold)) provided supporting evidence for the absence of an effect of any fraction on vascular dementia, with the sole exception of triglycerides. Raised triglycerides were associated with an increased incidence of vascular dementia across two previously unanalysed cohorts though there is the potential for uncontrolled confounding due to the limited information on covariates available. With respect to evidence on treatments for hypertriglyceridemia, examination of fibrates in the CPRD provided very weak evidence of an association with vascular dementia.

@@

```{r tgVadBIAMA, include = F}
```

@@

#### Case study #1: effect of LDL-c at midlife on Alzheimer's disease

From the data sources identified in Section \@ref(tri-data-sources), I identified 18 results relevant to my first causal question of interest. Most (n=17) were identified via the systematic review,`r ldl_ad_citations` while Chapter \@ref(cprd-heading) provided an additional estimate from the CPRD analysis.  The extent, type and direction of predicted bias for each result can be seen in the bias-direction plot presented in Figure \@ref(fig:ldlAdBiasDirection), stratified by study design.
